### 2014-12-01 paper/Content/01introduction.tex
added initial table of contents. however, with rough titles.

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....

\IEEEPARstart{D}{ieses} Template gilt f\"ur IEEE journal papers, die in \LaTeX\ mit der Klasse
IEEEtran.cls geschrieben werden.
% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)
Vielleicht k\"onnen wir ja ein sch\"ones Paper damit erstellen und uns so schon einen Pluspunkt bei der Abgabe abholen. Ich bin mir sicher, dass der Prof. Zimmermann auf sch\"ones Aussehen steht.

	\subsection{Android}

	\subsection{MVC \& MVP}

	\subsection{Problems in Android Development}

	\subsection{Motivation: MVP for Android}

	\subsection{Research Questions}

\section{Research}

\IEEEPARstart{I}{n} this chapter, the questions are answered by research. Blablablabalabalaba.

	\subsection{Android's Building Blocks}

	\subsection{Problems in Android Development (detailed)}

	\subsection{MVP's two flavors: Passive View \& Supervising Controller}

	\subsection{Approach to solving the problem}