### 2014-12-01 paper/Content/03conclusion.tex

\section{Conclusion}

\IEEEPARstart{U}{sing} this approach is the future of Android and makes everything better. That is because we are such cool guys and we'll become scientists in the next two years. Here are other good reasons:

	\subsection{Outlook of our approach}

	\subsection{Other people active in this context}