### 2014-12-02 paper/Content/sources.tex

\begin{thebibliography}{1}

\bibitem{GartnerAndroid2014}
Gartner, \emph{Gartner Says Worldwide Traditional PC, Tablet, Ultramobile and 
Mobile Phone Shipments Are On Pace to Grow 6.9 Percent in 2014}, March 2014, 
[Online]. http://www.gartner.com/newsroom/id/2692318 (retrieved 02-Dec-2014)

\bibitem{AppBrainStats}
App Brain, \emph{Android Statistics -- Number of Android applications}. 
[Online].
http://www.appbrain.com/stats/number-of-android-apps (retrieved 02-Dec-2014)

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

\todo[inline]{
Citation goes here. No BibTeX because it would be overkill, only disadvantage: 
We need to add sources in the right citation style manually.
}