### 2015-01-10 paper/Content/sources.tex
Added Fowler; added MVP UML-like graphic

\begin{thebibliography}{1}

\bibitem{PassiveMVC}
K. Sokolova, M. Lemercier, and L. Garcia: \emph{Android Passive MVC: a Novel Architecture Model for Android Application Development.} In: PATTERNS 2013, The Fifth International Conferences on Pervasive Patterns and Applications, p. 7-12, 2013.

\bibitem{GartnerAndroid2014}
Gartner: \emph{Gartner Says Worldwide Traditional PC, Tablet, Ultramobile and 
Mobile Phone Shipments Are On Pace to Grow 6.9 Percent in 2014}, March 2014, 
[Online]. http://www.gartner.com/newsroom/id/2692318 (retrieved 02-Dec-2014)

\bibitem{AppBrainStats}
App Brain: \emph{Android Statistics -- Number of Android applications}. 
[Online].
http://www.appbrain.com/stats/number-of-android-apps (retrieved 02-Dec-2014)

\bibitem{AndroidDeveloperCollection}
J. Steele, N. To, S. Conder, and L. Darcey: \emph{The Android Developer's Collection.} Addison-Wesley Professional, December 2011.

\bibitem{SDKActivities}
Google Inc., \emph{Android Developers -- Activities},
[Online]. http://developer.android.com/guide/components/activities.html (retrieved 05-Jan-2015)

\bibitem{SDKFragments}
Google Inc., \emph{Android Developers -- Fragments},
[Online]. http://developer.android.com/guide/components/fragments.html(retrieved 05-Jan-2015)

\bibitem{BallOfMud}
B. Foote and J. Yoder: \emph{Big Ball of Mud.} Addison-Wesley, 1997.

\bibitem{TaligentMVP}
M. Potel: \emph{MVP: Model-View-Presenter the taligent programming model for C++ and Java}, Taligent Inc., 1996.

\bibitem{GangOfFour}
E. Gamma, R. Helm, R. Johnson, and J. Vlissides: \emph{Design Patterns: Elements of Reusable Object-Oriented Software,} 1\textsuperscript{st} ed. Addison-Wesley Professional, November 1994.

\bibitem{FowlerMvpRetirementNote}
Fowler, Martin: \emph{Retirement note for Model View Presenter Pattern},
[Online]. http://martinfowler.com/eaaDev/ModelViewPresenter.html (retrieved 10-Jan-2015)

\bibitem{FowlerSupervisingController}
Fowler, Martin: \emph{Supervising Controller},
[Online]. http://martinfowler.com/eaaDev/SupervisingPresenter.html (retrieved 10-Jan-2015)

\bibitem{FowlerPassiveView}
Fowler, Martin: \emph{Passive View},
[Online]. http://martinfowler.com/eaaDev/PassiveScreen.html (retrieved 10-Jan-2015)

%\bibitem{}
%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}