### 2015-01-13 paper/Content/abstract.tex
First Abstract Idea

\begin{abstract}
%\boldmath
One of the most commonly used operating system in mobile industry is \emph{Android}. This paper is about development of \emph{Android} applications and describes how an Model-View-Presenter architecture can be applied to such. An application prototype was developed to support this paper.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway.

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Android,  Application Architecture, Model-View-Presenter (MVP), App Prototype.
\end{IEEEkeywords}